---
title: "TidyVerse"
author: "Weijian Cai"
date: "May 4, 2019"
output: html_document
---

###The tidyverse is a powerfull collecion fo R packages that are great data tools for data transformation and data visualization. 
###I will cover the ggplot2 packages to show how to make different graphs like line plot, bar plot and  scater plot.And demostate how to custom plots by adding mutiple layers beyond. 
###The data source I use is from kaggle dataset.

load the core tidyverse package
```{r}
library(tidyverse)
```

load the dataset from a csv file into R.
```{r}
url<-"https://raw.githubusercontent.com/Weicaidata/TidyVerse/master/athlete_events.csv"
url2<-"https://raw.githubusercontent.com/Weicaidata/TidyVerse/master/noc_regions.csv"


df<-read_csv(url)
region<-read_csv(url2)
head(df)
head(region)
```
###line plot 
line plot connects the dots in order of the variable present on the x-axis.I will show you how to Create a ggplot line plot below.

####number of athletes by Year and Sex

first we use dplyr to performance data transfermation.
```{r}
counts_sex <- df %>% 
              group_by(Year, Sex) %>%
              summarize(Athletes = length(unique(ID)))
counts_sex$Year <- as.integer(counts_sex$Year)
```


basic arguement to create line plot
```{r}
ggplot(counts_sex, aes(x=Year, y=Athletes, group=Sex, color=Sex)) +  #assign color based on Sex
  geom_point(size=2) +                          # to create a line plot
  geom_line()  +
  scale_color_manual(values=c("red","blue")) +  # modify color of each lines
  labs(title = "Number of male and female") +   # add title above the graph
  theme(plot.title = element_text(hjust = 0.5)) #change theme
```

####Number of Countries participated over time.

use dply package - filter, arrange, mutate, summarize function to manipulate dataset.
```{r}
counts <- df %>%
          filter(Team != "Unknown") %>%
          group_by(Year) %>%
          summarize(Nations = length(unique(Team)))
```

another example of line plot
```{r}
ggplot(counts, aes(x=Year, y=Nations)) +   #initiate plotting
  geom_point(size=2) +                     #basic line plot parameter
  geom_line() + 
  xlab("")
```
### Bar plot

####medal winners by country
```{r}
medals<- df%>%filter(!is.na(Medal))

medals<- medals%>%left_join(region,by=c("NOC"="NOC"))
medals$notes = NULL
medals<-medals%>%
        group_by(region)%>%
        summarize(total=n())%>%
        arrange(desc(total))%>%
        head(n=20)

medals$region <- factor(medals$region,levels = medals$region[order(medals$total)])
```

Bar plot example explanation
```{r}
ggplot(medals,aes(region,total,color=region,fill=region)) +             #initate ggplot
  geom_bar(position = "stack",  width =.6,stat="identity") +            #to draw bar geom
  coord_flip()+                                                         # to make a horizontal bar chart
  geom_text(aes(label=total,hjust=-.03,  colour="black"),size=3)+       # add text on each bars
  theme(axis.line = element_line(color = "yellow",size=1))+             # change theme
      theme(panel.background=element_blank())+             
      scale_x_discrete() +
    xlab("Country")+ylab("Medals")+                                     # change x and y axis lable
  theme(legend.position = "none",
                     axis.text = element_text(size = 8,face="bold"),
        plot.title = element_text(size=16,face = "bold")) +  
  ggtitle("Top medal by Country")                                      # add a title
```




